### 2014-08-29 paper/paper.Rmd
tweaked albania

```{r age, fig.width=10, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE}
student2012.sub <- student2012[, c(1, 6, 12, 14, 501, 541, 546, 634)]
student2012.sub.age <- summarise(group_by(student2012.sub, name), 
                  math4 = mean(PV1MATH[ST06Q01 == 4], na.rm = T), 
                  math5 = mean(PV1MATH[ST06Q01 == 5], na.rm = T), 
                  math6 = mean(PV1MATH[ST06Q01 == 6], na.rm = T), 
                  math7 = mean(PV1MATH[ST06Q01 == 7], na.rm = T), 
                  math8 = mean(PV1MATH[ST06Q01 >  7], na.rm = T))
student2012.sub.age.m <- melt(student2012.sub.age)
student2012.sub.age.m$variable <- substr(student2012.sub.age.m$variable, 5, 5)
student2012.sub.age.m$variable <- as.numeric(student2012.sub.age.m$variable)
student2012.sub.agemedian <- summarise(group_by(student2012.sub, name), medianstart = median(ST06Q01, na.rm=T))
orderedbymedianage <- student2012.sub.agemedian$name[order(student2012.sub.agemedian$medianstart)]
student2012.sub.age.m.median <- left_join(student2012.sub.age.m, student2012.sub.agemedian)
student2012.sub.age.m.median$medianstart <- factor(student2012.sub.age.m.median$medianstart)
student2012.sub.age.m.median$name <- factor(student2012.sub.age.m.median$name, levels=orderedbymedianage)
qplot(variable, value, data=student2012.sub.age.m.median, xlab="Age at start", ylab="Math Score", 
      geom=c("point","smooth"), se=F, group=name, colour=medianstart) + 
 facet_wrap(~name, ncol=8) + theme_bw() + theme(legend.position="none")
# ggsave("school-start.pdf", width=14, height=14)
#student2012.sub.age <- summarise(group_by(student2012.sub, name), 
#                  read4 = mean(PV1READ[ST06Q01 == 4], na.rm = T), 
#                  read5 = mean(PV1READ[ST06Q01 == 5], na.rm = T), 
#                  read6 = mean(PV1READ[ST06Q01 == 6], na.rm = T), 
#                  read7 = mean(PV1READ[ST06Q01 == 7], na.rm = T), 
#                  read8 = mean(PV1READ[ST06Q01 >  7], na.rm = T))
#student2012.sub.age.m <- melt(student2012.sub.age)
#student2012.sub.age.m$variable <- substr(student2012.sub.age.m$variable, 5, 5)
#student2012.sub.age.m$variable <- as.numeric(student2012.sub.age.m$variable)
#qplot(variable, value, data=student2012.sub.age.m, xlab="Age at start", ylab="Reading Score", 
#      geom=c("point","smooth"), se=F, group=name) + 
# facet_wrap(~name, ncol=8) + theme_bw() 
#student2012.sub.age <- summarise(group_by(student2012.sub, name), 
#                  scie4 = mean(PV1SCIE[ST06Q01 == 4], na.rm = T), 
#                  scie5 = mean(PV1SCIE[ST06Q01 == 5], na.rm = T), 
#                  scie6 = mean(PV1SCIE[ST06Q01 == 6], na.rm = T), 
#                  scie7 = mean(PV1SCIE[ST06Q01 == 7], na.rm = T), 
#                  scie8 = mean(PV1SCIE[ST06Q01 >  7], na.rm = T))
#student2012.sub.age.m <- melt(student2012.sub.age)
#student2012.sub.age.m$variable <- substr(student2012.sub.age.m$variable, 5, 5)
#student2012.sub.age.m$variable <- as.numeric(student2012.sub.age.m$variable)
#qplot(variable, value, data=student2012.sub.age.m, xlab="Age at start", ylab="Science Score", 
#      geom=c("point","smooth"), se=F, group=name) + 
# facet_wrap(~name, ncol=8) + theme_bw() 
```

@@

```{r clean, echo=FALSE}
student2012$CNT = student2012$name
student2012$OECD <- as.character(student2012$OECD)
student2012[student2012$CNT == "Albania", ]$OECD = "Albania"
student2012$math = as.numeric(student2012[, "PV1MATH"])
student2012$read = as.numeric(student2012[, "PV1READ"])
student2012$sci = as.numeric(student2012[, "PV1SCIE"])

student2012.sub = student2012 %>% dplyr::select(CNT, SCHOOLID, math, read, sci, OUTHOURS, STIDSTD, ST04Q01, ST11Q01, ST11Q02, ST15Q01, ST19Q01, ST26Q01, ST26Q02, ST26Q03, ST26Q04, ST26Q05, ST26Q06, ST26Q07, ST26Q08, ST26Q09, ST26Q10, ST26Q11, ST26Q12, ST26Q13, ST26Q14, ST27Q01, ST27Q02, ST27Q03, ST27Q04, ST27Q05, ST28Q01, OECD)

@@

```{r albania-possessions, echo=FALSE}
df = melt(student2012.sub %>% group_by(CNT, numposs, OECD) %>% summarise(mmath = mean(math), msci = mean(sci), mread = mean(read)), id=c("CNT", "numposs", "OECD"))

df$group = paste0(df$CNT, df$variable)
qplot(x = numposs, y = value, data = df, facets = ~OECD, geom = "line", color = variable, group = group) + ylab("test scores") + xlab("number of possessions") + scale_colour_discrete(name = "Subject", labels=c("Math", "Science", "Reading"))
```

```{r truancyschool, echo=FALSE}
school2012$CNT = school2012$name
school2012.sub = school2012 %>% dplyr::select(CNT, SCHOOLID, SC22Q01)
student2012.sub = student2012 %>% dplyr::select(CNT, SCHOOLID, math, read, sci, OECD)
truancy.df = melt( student2012.sub %>% 
  group_by(CNT, SCHOOLID) %>% 
  left_join(school2012.sub, by = c("CNT", "SCHOOLID")) %>% 
  group_by(CNT, SC22Q01, OECD) %>% summarise(math = mean(math), read = mean(read), sci = mean(sci)), 
  id = c("CNT", "SC22Q01", "OECD")) %>% filter(!is.na(SC22Q01))
truancy.df$group = paste0(truancy.df$CNT, truancy.df$variable)

qplot(x = SC22Q01, y = value, color = variable, facets = ~OECD, data = truancy.df, group = group, geom = "line") + xlab("reported truancy") + ylab("test scores") + scale_colour_discrete(name = "Subject", labels=c("Math", "Science", "Reading"))
```

@@

data_small <- 
  student2012 %>%
  dplyr::select(OECD, CNT, OUTHOURS) %>%
  filter(!is.na(OUTHOURS))

ggplot(mapping = aes(x = OUTHOURS)) + 
  scale_x_continuous(name = "out of school study time (hours)", limits = c(0,60)) +
  scale_y_continuous(name = "density") +
  geom_histogram(
    mapping = aes(y = ..density.., fill = OECD), 
    binwidth = .width,
    data = data_small %>% filter(OECD == "Albania")
  ) + 
  scale_fill_manual(
    values = "black",
    breaks = "Albania",
    guide = guide_legend(title = "")
  ) + 
  geom_line(
    mapping = aes(group = CNT, color = OECD), 
    stat = "density", 
    alpha = 0.3,
    data = data_small %>% filter(OECD != "Albania")
  ) +
  scale_color_discrete(
    guide = guide_legend(
      title = "Other Countries",
      override.aes = list(alpha = 1)
    )

@@

```{r gendergap, echo=FALSE, warning = FALSE, fig.height = 10}
scores = student2012 %>%
  dplyr::select(CNT, math, read, ST04Q01, OECD) %>% 
  group_by(OECD, CNT, ST04Q01) %>%
  summarise(math = mean(math), read = mean(read))
male = scores %>% filter(ST04Q01 == "Male")
female = scores %>% filter(ST04Q01 == "Female")
scores.df = data.frame(OECD = male$OECD, CNT = male$CNT, math_difference = male$math - female$math, read_difference = male$read - female$read)

scores.df$CNT = factor(scores.df$CNT, levels = scores.df$CNT[order(scores.df$math_difference)])
scores.df$OECD = factor(scores.df$OECD, levels = scores.df$OECD[])

qplot(CNT, math_difference, data = scores.df, color = OECD, size = OECD) + coord_flip() +
  scale_y_continuous("math score gap: male - female scores", breaks=seq(-30, 30, 5)) +
  geom_hline(yintercept=0, colour="grey80") + scale_size_manual(values = c(4, 2.5, 2.5)) + xlab("") +  
  scale_colour_manual(values = c("#ca0020", "#f4a582", "#92c5de"))

scores.df$CNT = factor(scores.df$CNT, levels = scores.df$CNT[order(scores.df$read_difference)])
qplot(CNT, read_difference, data = scores.df, color = OECD, size = OECD) + coord_flip() +
  scale_y_continuous("read score gap: male - female scores") +
  geom_hline(yintercept=0, colour="grey80") + scale_size_manual(values = c(4, 2.5, 2.5)) + xlab("") +  
  scale_colour_manual(values = c("#ca0020", "#f4a582", "#92c5de"))
```

>>>>

```{r age, fig.width=10, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE}
student2012.sub <- student2012[, c(1, 6, 12, 14, 501, 541, 546, 634)]
student2012.sub.age <- summarise(group_by(student2012.sub, name), 
                  math4 = mean(PV1MATH[ST06Q01 == 4], na.rm = T), 
                  math5 = mean(PV1MATH[ST06Q01 == 5], na.rm = T), 
                  math6 = mean(PV1MATH[ST06Q01 == 6], na.rm = T), 
                  math7 = mean(PV1MATH[ST06Q01 == 7], na.rm = T), 
                  math8 = mean(PV1MATH[ST06Q01 >  7], na.rm = T))
student2012.sub.age.m <- melt(student2012.sub.age)
student2012.sub.age.m$variable <- substr(student2012.sub.age.m$variable, 5, 5)
student2012.sub.age.m$variable <- as.numeric(student2012.sub.age.m$variable)
student2012.sub.agemedian <- summarise(group_by(student2012.sub, name), medianstart = median(ST06Q01, na.rm=T))
orderedbymedianage <- student2012.sub.agemedian$name[order(student2012.sub.agemedian$medianstart)]
student2012.sub.age.m.median <- left_join(student2012.sub.age.m, student2012.sub.agemedian)
student2012.sub.age.m.median$medianstart <- factor(student2012.sub.age.m.median$medianstart)
student2012.sub.age.m.median$name <- factor(student2012.sub.age.m.median$name, levels=orderedbymedianage)
qplot(variable, value, data=student2012.sub.age.m.median, xlab="Age at start", ylab="Math Scores", 
      geom=c("point","smooth"), se=F, group=name, colour=medianstart) + 
 facet_wrap(~name, ncol=8) + theme_bw() + theme(legend.position="none")
# ggsave("school-start.pdf", width=14, height=14)
#student2012.sub.age <- summarise(group_by(student2012.sub, name), 
#                  read4 = mean(PV1READ[ST06Q01 == 4], na.rm = T), 
#                  read5 = mean(PV1READ[ST06Q01 == 5], na.rm = T), 
#                  read6 = mean(PV1READ[ST06Q01 == 6], na.rm = T), 
#                  read7 = mean(PV1READ[ST06Q01 == 7], na.rm = T), 
#                  read8 = mean(PV1READ[ST06Q01 >  7], na.rm = T))
#student2012.sub.age.m <- melt(student2012.sub.age)
#student2012.sub.age.m$variable <- substr(student2012.sub.age.m$variable, 5, 5)
#student2012.sub.age.m$variable <- as.numeric(student2012.sub.age.m$variable)
#qplot(variable, value, data=student2012.sub.age.m, xlab="Age at start", ylab="Reading Score", 
#      geom=c("point","smooth"), se=F, group=name) + 
# facet_wrap(~name, ncol=8) + theme_bw() 
#student2012.sub.age <- summarise(group_by(student2012.sub, name), 
#                  scie4 = mean(PV1SCIE[ST06Q01 == 4], na.rm = T), 
#                  scie5 = mean(PV1SCIE[ST06Q01 == 5], na.rm = T), 
#                  scie6 = mean(PV1SCIE[ST06Q01 == 6], na.rm = T), 
#                  scie7 = mean(PV1SCIE[ST06Q01 == 7], na.rm = T), 
#                  scie8 = mean(PV1SCIE[ST06Q01 >  7], na.rm = T))
#student2012.sub.age.m <- melt(student2012.sub.age)
#student2012.sub.age.m$variable <- substr(student2012.sub.age.m$variable, 5, 5)
#student2012.sub.age.m$variable <- as.numeric(student2012.sub.age.m$variable)
#qplot(variable, value, data=student2012.sub.age.m, xlab="Age at start", ylab="Science Score", 
#      geom=c("point","smooth"), se=F, group=name) + 
# facet_wrap(~name, ncol=8) + theme_bw() 
```

@@

```{r clean, echo=FALSE}
#student2012$CNT = student2012$name
student2012$OECD <- as.character(student2012$OECD)
#student2012[student2012$name == "Albania", ]$OECD = "Albania"
student2012$math = as.numeric(student2012[, "PV1MATH"])
student2012$read = as.numeric(student2012[, "PV1READ"])
student2012$sci = as.numeric(student2012[, "PV1SCIE"])

student2012.sub = student2012 %>% dplyr::select(name, SCHOOLID, math, read, sci, OUTHOURS, STIDSTD, ST04Q01, ST11Q01, ST11Q02, ST15Q01, ST19Q01, ST26Q01, ST26Q02, ST26Q03, ST26Q04, ST26Q05, ST26Q06, ST26Q07, ST26Q08, ST26Q09, ST26Q10, ST26Q11, ST26Q12, ST26Q13, ST26Q14, ST27Q01, ST27Q02, ST27Q03, ST27Q04, ST27Q05, ST28Q01, OECD)

@@

```{r albania-possessions, fig.width=10, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE}
#df = melt(student2012.sub %>% group_by(name, numposs, OECD) %>% summarise(mmath = mean(math), msci = mean(sci), mread = mean(read)), id=c("name", "numposs", "OECD"))
df = melt(student2012.sub %>% group_by(name, numposs, OECD) %>% summarise(mmath = mean(math)), id=c("name", "numposs", "OECD"))

#df$group = paste0(df$name, df$variable)
#qplot(x = numposs, y = value, data = df, facets = ~OECD, geom = "line", color = variable, group = group) + ylab("test scores") + xlab("number of possessions") + scale_colour_discrete(name = "Subject", labels=c("Math", "Science", "Reading"))
qplot(x = numposs, y = value, data = df, facets = ~name, geom = "line") + ylab("Math Scores") + xlab("Number of Possessions") + theme_bw()
```

```{r truancyschool, fig.width=10, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE}
#school2012$name = school2012$name
school2012.sub = school2012 %>% dplyr::select(name, SCHOOLID, SC22Q01)
student2012.sub = student2012 %>% dplyr::select(name, SCHOOLID, math, read, sci, OECD)
#truancy.df = melt( student2012.sub %>% 
#  group_by(name, SCHOOLID) %>% 
#  left_join(school2012.sub, by = c("name", "SCHOOLID")) %>% 
#  group_by(name, SC22Q01, OECD) %>% summarise(math = mean(math), read = mean(read), sci = mean(sci)), 
#  id = c("name", "SC22Q01", "OECD")) %>% filter(!is.na(SC22Q01))
truancy.df = melt( student2012.sub %>% 
  group_by(name, SCHOOLID) %>% 
  left_join(school2012.sub, by = c("name", "SCHOOLID")) %>% 
  group_by(name, SC22Q01, OECD) %>% summarise(math = mean(math)), 
  id = c("name", "SC22Q01", "OECD")) %>% filter(!is.na(SC22Q01))
#truancy.df$group = paste0(truancy.df$name, truancy.df$variable)

#qplot(x = SC22Q01, y = value, color = variable, facets = ~name, data = truancy.df, group = group, geom = "line") + xlab("Reported Truancy") + ylab("Test Scores") + scale_colour_discrete(name = "Subject", labels=c("Math", "Science", "Reading"))
qplot(x = SC22Q01, y = value, facets = ~name, data = truancy.df, geom = "line", group=name) + xlab("Reported Truancy") + ylab("Math Scores") + theme_bw()
```

@@

data_small <- 
  student2012 %>%
  dplyr::select(OECD, name, OUTHOURS) %>%
  filter(!is.na(OUTHOURS))

ggplot(mapping = aes(x = OUTHOURS)) + 
  scale_x_continuous(name = "out of school study time (hours)", limits = c(0,60)) +
  scale_y_continuous(name = "density") +
  geom_histogram(
    mapping = aes(y = ..density.., fill = OECD), 
    binwidth = .width,
    data = data_small %>% filter(OECD == "Albania")
  ) + 
  scale_fill_manual(
    values = "black",
    breaks = "Albania",
    guide = guide_legend(title = "")
  ) + 
  geom_line(
    mapping = aes(group = name, color = OECD), 
    stat = "density", 
    alpha = 0.3,
    data = data_small %>% filter(OECD != "Albania")
  ) +
  scale_color_discrete(
    guide = guide_legend(
      title = "Other Countries",
      override.aes = list(alpha = 1)
    )

@@
